 

### 2021-05-05 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

**Stability of $T_{leaf}$ varies with fluctuation in biophysical conditions**
Biophysical environment fluctuation dictates how stable $T_{leaf}$ is, where stable biophysical conditions, for eg. low wind speed and consistent $T_{air}$ can prolong $T_{leaf}$, consistently reducing stomatal conductance and photosynthesis while increasing $T_{leaf}$, and therefore reducing chances for leaf recovery. However, increase in biophysical fluctuations make $T_{leaf}$ unstable as seen in upper canopies where with available water unstable $T_{leaf}$ can be more manageable to peak irradiance and $T_{air}$ than in stable buffered canopy regions [**REF**]

@@

*Among species that can be deciduous, greater proportion of deciduous individuals in larger size classes (Condit et al. 2000). *
*Spring and fall leaf phenology in temperate deciduous forests (Augsburger).*

@@

Across tree-size ontogeny, canopy and understory trees of same species in tropical, temperate, and conifer forests show fundamental light adapted leaf traits in their biophysical environment, as observed and similar to within-species gradient [@rozendaalPlasticityLeafTraits2006; @thomasPhotosyntheticDifferencesSaplings2002; @houterOntogeneticChangesLeaf2012]. Isolating ontogeny, under constant light conditions of open-grown canopy and understory trees, for example in large gaps, differences in traits showed significantly greater LMA, photosynthetic capcity per area, and stomatal conductance in taller canopy trees compared to saplings [@thomasPhotosyntheticDifferencesSaplings2002; @houterOntogeneticChangesLeaf2012]. Across  multi-species canopy and understory, while fundamental adaptations to light and water are observed, their pasticity to light differs along with inherent traits such as leaf shape, size and width, stomtal conductance levels and therefore influence leaf temperatures.As a result, some species might be more susceptible to heat stress at a vertical gradient than others [@fauset_differences_2018].

>>>>

**Stability of $T_{leaf}$ varies with fluctuation in biophysical conditions**
Biophysical environment fluctuation dictates how stable $T_{leaf}$ is, where stable biophysical conditions, for eg. low wind speed and consistent $T_{air}$ can prolong $T_{leaf}$, consistently reducing stomatal conductance and photosynthesis while increasing $T_{leaf}$, and therefore reducing chances for leaf recovery. However, increase in biophysical fluctuations, for eg. wind speed, cloud mediated irradiance fluctuation and $T_{air}$  make $T_{leaf}$ unstable as seen in upper canopies where with available water unstable $T_{leaf}$ can be more manageable at peak irradiance and $T_{air}$ than in stable buffered canopy regions [**REF**] ~this understanding needs more help from co-authors, should this be included? (thermal time constant?)~

@@

Deciduousness as a leaf trait is an adaptation to water stress therefore can influence biophysical gradient and leaf temperature distribution during leaf-loss [@rey-sanchez_spatial_2016]. *Among species that can be deciduous, greater proportion of deciduous individuals in larger size classes that also tend to lose their leaves than understory species that are seldom deciduous in the tropics. Deciduous species are also more prominent in drier sites compared to wet sites [@conditQuantifyingDeciduousnessTropical2000b]*

*Spring and fall leaf phenology in temperate deciduous forests (Augsburger).* Leaf emergence in deciduous forests is increasing delayed with height-ontogeny where understory species leaf-out earlier than canopy trees [Vitasse et al. 2013; @seiwaChangesLeafPhenology1999].

@@

Across tree-size ontogeny, canopy and understory trees of same species in tropical, temperate, and conifer forests show fundamental light adapted leaf traits in their biophysical environment, as observed and similar to within-species gradient [@rozendaalPlasticityLeafTraits2006; @thomasPhotosyntheticDifferencesSaplings2002; @houterOntogeneticChangesLeaf2012]. Isolating ontogeny, under constant light conditions of open-grown canopy and understory trees, for example in large gaps, differences in traits show significantly greater LMA, photosynthetic capcity per area, and stomatal conductance in taller canopy trees compared to saplings [@thomasPhotosyntheticDifferencesSaplings2002; @houterOntogeneticChangesLeaf2012]. Across  multi-species canopy and understory, while fundamental adaptations to light and water are observed, their pasticity to light differs along with inherent traits such as leaf shape, size and width, stomtal conductance levels and therefore influence leaf temperatures.As a result, some species might be more susceptible to heat stress at a vertical gradient than others [@fauset_differences_2018].
